%-------------------------------------------------------------------------------
% first_start
%-------------------------------------------------------------------------------
%
% \file        first_start.tex
% \library     Documents
% \author      Chris Ahlstrom
% \date        2015-11-01
% \update      2023-12-04
% \version     $Revision$
% \license     $XPC_GPL_LICENSE$
%
%     This document provides LaTeX documentation for Seq66.
%
%-------------------------------------------------------------------------------

\section{Let's Go!}
\label{sec:introduction_lets_go}

   \textsl{Seq66} requires ALSA or JACK on Linux;
   it does not support
   \index{pipewire}
   Pipewire.
   It might work with a Pipewire-JACK setup, but that usage has not been
   tested.
   Another aspect to look into is interactions with
   PulseAudio-pw.
   On Windows, \textsl{Seq66} uses the \textsl{MultiMedia API} with a wrapper
   that is our modified version of \textsl{PortMIDI} \cite{portmidi}.
   Theoretically that library should support
   \textsl{Mac OSX}, but we have no Mac to build and test with.
   This manual assumes that one is running \textsl{Seq66} on Linux, for the
   most part.
   But see \sectionref{sec:windows}.

   For the first run, make sure all desired MIDI devices are plugged in and
   that all desired software synthesizers are running.
   Only ports present at startup will appear; this version of
   \textsl{Seq66} does not
   support automatic handling of port changes at run-time.
   In addition, port-mapping (see \sectionref{sec:port_mapping}) is
   enabled by default; this makes the port setup more flexible..

   Start \textsl{Seq66}, verify that it comes up, and then exit it
   immediately.
   This first start creates the configuration files in

   \begin{verbatim}
      /home/user/.config/seq66/}                      (Linux)
      C:/Users/your\_user\_name/AppData/Local/seq66}  (Windows)
   \end{verbatim}

   If a session manager (e.g. \textsl{NSM})
   is used, the configuration directory is determined by
   the session manager, and is also called a "session" directory..
   The main configuration file is
   \texttt{qseq66.rc} (Linux) or
   \texttt{qpseq66.rc} (Windows).
   The '.rc' file contains ports and port-mapping for the devices
   initially on the system.
   It also specifies the names of a number of other configuration files.

   Now start \textsl{Seq66} to use JACK or ALSA for MIDI.
   For USB MIDI devices on JACK, the \texttt{a2jmidid} program might
   need to be running
   (see \sectionref{sec:jack}, for details.)
   On \textsl{Windows}, just run \texttt{qpseq66.exe};
   see \sectionref{sec:windows}.

   For better trouble-shooting, run it via command-line in
   a terminal window, at first and provide a MIDI file, to
   see what messages might appear.
   Also note that a log file is created; by default message will be
   written to it.
   The port settings will depend on your system.
   On our system, the synthesizer (\textsl{Yoshimi}) comes up on MIDI buss 5.
   (On \textsl{Windows}, buss 0 is the "MIDI Mapper", while buss 1 is the
   built-in wavetable synthesizer, which is normally under control of buss 0.)
   The \texttt{-{}-buss} option remaps all events to the desired buss for
   the current run:

   \begin{verbatim}
      $ qseq66 --alsa --buss 5 /usr/share/seq66/midi/b4uacuse-gm-patchless.midi
      $ qseq66 --jack --buss 5 /usr/share/seq66/midi/b4uacuse-gm-patchless.midi
   \end{verbatim}

   The buss-override setting can also be made via the port drop-down control
   in the main window. It will modify the MIDI file.
   The "data" directory is a directory created upon installation of the
   application:

   \begin{verbatim}
      /usr/share/seq66-0.99/                          (Linux)
      C:/Program Files/Seq66/data/                    (Windows)
   \end{verbatim}

   It contains sample MIDI files and configuration files.
   Some of the files in those directories apply to both operating systems, so
   be sure to look at them.
   The user configration files are stored in the user's "home" area:

   \begin{verbatim}
      /home/username/.config/seq66/qseq66.*           (Linux)
      C:/Users/username/AppData/Local/seq66           (Windows)
      ~/NSM Sessions/sessname/seq66.nXXXX/config      (NSM)
   \end{verbatim}

   The configuration files in the "home" directory
   are created after the first run of \textsl{Seq66}.
   If all is well, the "play" button will start playback and the tune sounds.
   If not, look into \textbf{Edit / Preferences} or the configuration files.
   Don't change the configuration files while \textsl{Seq66} is running, as
   they are generally re-saved at application exit.

   \textbf{Bug?}
      The first-start change of input port settings might not get saved
      to the 'rc' file. The workaround is to exit and rerun
      \textsl{Seq66}, and make the setting again.
      Mysterious.

\subsection{Device Changes}
\label{subsec:introduction_device_changes}

   After running \textsl{Seq66} and getting it to work, one might
   either add new MIDI devices and software synthesizers, or remove
   them.
   If port-mapping is enabled (the default, and recommended), then
   one can go to
   \textbf{Edit / Preferences / MIDI Clock} and
   click the \textbf{Make Maps} button.
   This action stores (again) the current set of MIDI devices.
   Restart \textsl{Seq66} and the new set of devices should appear.

   \textsl{Seq66} will also detect at startup if there are unavailable
   ports or if there are now more real MIDI ports than mapped ports.
   It will prompt for a potential remapping and restart.
   Click that button if feeling lucky, otherwise
   exit and fix the 'rc' file with a text editor.

   \textbf{Warning}:
   If one has hand-tailored the port maps to represent all possible devices one
   has, the remap or restart buttons might \textbf{remove}
   some of those devices.
   Always keep a \textbf{backup}
   of a known-good 'rc' file in a safe place.

   In version 2 of \textsl{Seq66}, a year or two from now, we hope to make
   the adjustments automatic.

\subsection{Ports in Seq66 MIDI Files}
\label{subsec:introduction_ports}

   When opening a pre-existing \textsl{Seq66} which contains patterns with
   output port numbers that don't exist on the current system, a prompt
   will appear with option to remap-and-restart \textsl{Seq66}.
   However, what normally needs to be done is to change the
   port numbers to the numbers of existing ports, and then save the MIDI file.
   One way to do this is to use the "buss-override" feature
   (see \sectionref{subsubsec:introduction_sets_buss_override}),
   which affects all patterns shown on the Live grid.

\subsection{Windows}
\label{subsec:introduction_windows}

% TODO: talk about LoopMIDI

   Details about \textsl{Windows} are in the file
   \texttt{C:/Program Files/Seq66/data/midi/readme.windows}
   and in \sectionref{sec:windows}.
   The first-start depends on if it is a bare \textsl{Windows}
   install, or if supplemental MIDI support such as 
   the 
   \textsl{CoolSoft MIDIMapper} and 
   \textsl{VirtualMIDISynth} have been installed.
   If they haven't, the default MIDI setup will contain no MIDI inputs,
   and two MIDI outputs, the \textsl{MIDI Mapper} and the
   \textsl{Microsoft GS Wavetable} synthesizer.
   The former grabs control of the latter, making it unavailable!
   \textsl{Seq66} detects this situation and enables only the MIDI Mapper.

   Therefore, run \textsl{Seq66} for the first time, then exit it and
   restart it. One should then be able to play MIDI files to the MIDI Mapper,
   if nothing else.
   We hope to get some input from
   \textsl{Windows} users, as our main usage is \textsl{Linux}
   and we run \textsl{Windows} only in an old virtual machine.

   As for virtual ports in \textsl{Windows}, a useful too
   is \textsl{LoopMIDI} (\cite{loopmidi}), as noted in the section
   on \textsl{Windows}.

%-------------------------------------------------------------------------------
% vim: ts=3 sw=3 et ft=tex
%-------------------------------------------------------------------------------
