%-------------------------------------------------------------------------------
% references
%-------------------------------------------------------------------------------
%
% \file        references.tex
% \library     Documents
% \author      Chris Ahlstrom
% \date        2015-08-31
% \update      2023-09-05
% \version     $Revision$
% \license     $XPC_GPL_LICENSE$
%
%     Provides the References section of the Seq66 manual. Rather
%     than use the bibtex package, our small set of references uses a
%     simpler method.
%
% Potential additional references:
%
%     cakewalk
%     WRK
%
%-------------------------------------------------------------------------------

\section{References}
\label{sec:references}

   The \textsl{Seq66} reference list.

{\RaggedRight
\begin{thebibliography}{99}

   \bibitem{alsa}
   ALSA team.
   \emph{Advanced Linux Sound Architecture (ALSA) project homepage.}
   \url{http://www.alsa-project.org/}.
   ALSA tools through version 1.0.29.
   2015.

   \bibitem{alsathru}
   superuser.com.
   \emph{How do I increase the number of MIDI through ports in ALSA?}
   \url{https://superuser.com/questions/973973/how-do-i-increase-the-number-of-midi-through-ports-in-alsa}
   2015.

   \bibitem{agordejo}
   laborejo.
   \emph{Agordejo Session Manager, part of the Laborejo Software Suite}
   \url{https://www.laborejo.org/agordejo/},
   \url{http://git.laborejo.org/lss/agordejo.git}
   2021.

   \bibitem{combine}
   Jay Capela Music.
   \emph{"Combine": A Seq24 Demonstration.}
   \url{https://www.youtube.com/watch?v=fUiXbVT0bJQ}.
   2010.

   \bibitem{forks}
   Various.
   \emph{Forks of Seq66.}
   \url{https://github.com/sivecj/seq66}.
   2021.

   \bibitem{jack}
   JACK team.
   \emph{JACK Audio Connection Kit.}
   \url{http://jackaudio.org/}.
   2015.

   \bibitem{kepler34}
   Oli Kester.
   \emph{Kepler34: Seq24 for the 2010s.}
   \url{https://github.com/oli-kester/kepler34}.
   2010-2016.

   \bibitem{lau}
   Linux Audio Users.
   \emph{Unofficial Linux audio users IRC channel.}
   \url{http://kiwiirc.com/nextclient/#irc://irc.freenode.net/#lau}.
   2021.

   \bibitem{layk}
   Lassi Ylikojola.
   \emph{Many demo videos of Sequencer64.}
   \url{https://www.youtube.com/watch?v=YStYVjFv1TM},
   \url{https://www.youtube.com/watch?v=GBlEP8Ffqss},
   \url{https://www.youtube.com/watch?v=4gG8SvJxJkA&t=28s},
   \url{https://www.youtube.com/watch?v=n4Z4WPK6FpA}.
   2010-2017.

   \bibitem{libreav}
   Author unknown.
   \emph{A description of Seq66.}
   \url{https://libreav.org/software/seq66}.
   2019.

   \bibitem{loopmidi}
   Tobias Erichsen.
   \emph{Private stuff \& software for audio, midi and more}
   \url{https://www.tobias-erichsen.de/software.html}.
   2022.

   \bibitem{midihowto}
   Phil Kerr.
   \emph{The Linux MIDI-HOWTO.}
   \url{https://tldp.org/HOWTO/MIDI-HOWTO.html}.
   2002.

   \bibitem{midilinux}
   Murks.
   \emph{ALSA and JACK MIDI explained (by a dummy for dummies).}
   \url{https://freeshell.de/~murks/posts/ALSA_and_JACK_MIDI_explained_(by_a_dummy_for_dummies)/}.
   2010.

   \bibitem{midimapper}
   Coolsoft.
   \emph{Coolsoft MIDIMapper (Windows).}
   \url{https://coolsoft.altervista.org/en/midimapper}.
   2018.

   \bibitem{midiox}
   Jamie O'Connell and Jerry Jorgenrud.
   \emph{MIDI OX and other related software (Windows).}
   \url{http://www.midiox.com/}.
   2017.

   \bibitem{midisynth}
   Coolsoft.
   \emph{Coolsoft VirtualMIDISynth (Windows).}
   \url{https://coolsoft.altervista.org/en/virtualmidisynth}.
   2018.

   \bibitem{midicontrol}
   linuxaudio.org.
   \emph{seq24: toggle sequences with a MIDI controller.}
   \url{http://wiki.linuxaudio.org/wiki/seq24togglemiditutorial}.
   2013.

   \bibitem{midicontroltable}
   midi.org.
   \emph{Summary of MIDI Messages.}
   \url{https://www.midi.org/specifications/item/table-1-summary-of-midi-message#2}.
   Year unknown.

   \bibitem{midicvt}
   Chris Ahlstrom.
   \emph{Extension of midicomp/midi2text to convert between MIDI and ASCII
      text format.}
   \url{https://github.com/ahlstromcj/midicvt}.
   2015-2016.

   \bibitem{nanobasket}
   Roy Vegard.
   \emph{Configurator software for the Korg nanoSERIES of MIDI controllers.}
   \url{https://github.com/royvegard/Nano-Basket}.
   Warning: this code may require an earlier version of Python than present in
   one's Linux distro or in Windows.
   2015.

   \bibitem{nsm}
   JACK Audio.
   \emph{New Session Manager, an offshot of Non Session Manager.}
   \url{https://github.com/jackaudio/new-session-manager}
   2021.

   \bibitem{osc}
   Stephen Sinclair
   \emph{Open Sound Control (library).}
   \url{https://github.com/radarsat1/liblo}
   2019.

   \bibitem{patchbay}
   Simon W. Fielding.
   \emph{QjackCtl and the Patchbay.}
   \url{https://www.rncbc.org/drupal/node/76}.
   2008.

   \bibitem{piseq}
   sferamusic
   \emph{HOW-TO: DIY hybrid Sequencer / MIDI USB Hub ....}
   \url{https://www.rncbc.org/drupal/node/7://steemit.com/music/@sferamusic/how-to-diy-hybrid-sequencer-midi-usb-hub-aka-piseq-using-raspberry-pi-and-external-midi-controllers-part-1}.
   2017.

   \bibitem{portmidi}
   PortMedia team.
   \emph{Platform Independent Library for MIDI I/O.}
   \url{http://portmedia.sourceforge.net/}.
   2010.

   \bibitem{ppqncalculator}
   Benjamin Robert Tubb
   \emph{MIDI PPQN Duration Calculator.}
   \url{https://demonstrations.wolfram.com/MIDIPPQNDurationCalculator/}.
   2011.

   \bibitem{qjackctl}
   \emph{QjackCtl.}
   \url{https://sourceforge.net/projects/qjackctl/}.
   \url{https://qjackctl.sourceforge.io/qjackctl-index.html}.
   \url{https://git.code.sf.net/p/qjackctl/code qjackctl-git}.
   2021.

   \bibitem{raysession}
   Houston4444.
   \emph{RaySession Session Manager}
   \url{https://raysession.tuxfamily.org/en/}
   \url{https://github.com/Houston4444/RaySession}
   2021.

   \bibitem{rtmidi}
   Gary P. Scavone.
   \emph{The RtMIDI Tutorial.}
   \url{https://www.music.mcgill.ca/~gary/rtmidi/}.
   2016.

   \bibitem{seq24}
   Seq24 Team.
   \emph{The home site for the Seq24 looping sequencer.}
   \url{http://www.filter24.org/seq24/download.html}.
   2010.

   \bibitem{seq24launchpad}
   Seq24 Team.
   \emph{The home site for the Seq66 looping sequencer.}
   \url{https://edge/launchpad.net/seq24}.
   2016.

   \bibitem{seq24launchpadmapper}
   Excds.
   \emph{A simple mapping for toggling the LEDs on the Novation launchpad
   together with seq24.}
   \url{https://github.com/Excds/seq24-launchpad-mapper}.
   2013.

   \bibitem{seq32}
   Stan Preston (stazed).
   \emph{The home site for the Seq32 looping sequencer.}
   \url{https://github.com/Stazed/seq32}.
   2016.

   \bibitem{seq66}
   Chris Ahlstrom.
   \emph{A reboot of the Seq24 project as "Seq66".}
   \url{https://github.com/ahlstromcj/seq66/}.
   2015-2021.

   \bibitem{seqtime}
   Unknown.
   \emph{Timing accuracy and resolution.}
   \url{http://midi.teragonaudio.com/tutr/seqtime.htm}.
   2000?.

   \bibitem{timidity}
   Timidity++ Team.
   \emph{Download site for Timidity++ source code.}
   \url{http://sourceforge.net/projects/timidity/}.
   2015.

   \bibitem{vmpk}
   VMPK Team.
   \emph{Virtual MIDI Piano Keyboard.}
   \url{http://vmpk.sourceforge.net/}.
   2015.

   \bibitem{windowsmidi}
   Donya Quick
   \emph{Working with MIDI on Windows}
   \url{http://donyaquick.com/midi-on-windows/#x1-120004.1}.
   2021.

   \bibitem{wootangent1}
   The Wootangent man.
   \emph{Seq24 Tutorial Videos, Part 1 and Part 2.}
   \url{http://wootangent.net/2010/10/linux-music-tutorial-seq24-part-1/}.
   \url{http://wootangent.net/2010/10/linux-music-tutorial-seq24-part-2/}.
   2010.

   \bibitem{youtubelinks}
   Various.
   \emph{Links to other usages of the seq** applications.}
   \url{https://www.youtube.com/watch?v=LjHPtlvjgnM}
      AndrewsVintageKeyboards: Yamaha TG100 with seq66.
   \url{https://www.youtube.com/watch?v=ESHKiHx09AA&t=78s}
      AndrewsVintageKeyboards: Yamaha TG100 with sequencer66.
   \url{https://www.youtube.com/watch?v=\_r7qXOo7kgg}
      Julian L: Seq64 demo.
   \url{https://www.youtube.com/watch?v=VCga2X4lJfg}
      Julian L: Seq64 demo 2.
\end{thebibliography}
}

%-------------------------------------------------------------------------------
% vim: ts=3 sw=3 et ft=tex
%-------------------------------------------------------------------------------
