%-------------------------------------------------------------------------------
% seq66_references
%-------------------------------------------------------------------------------
%
% \file        seq66_references.tex
% \library     Documents
% \author      Chris Ahlstrom
% \date        2015-08-31
% \update      2018-05-28
% \version     $Revision$
% \license     $XPC_GPL_LICENSE$
%
%     Provides the References section of the Sequencer66 manual. Rather
%     than use the bibtex package, our small set of references uses a
%     simpler method.
%
%-------------------------------------------------------------------------------

\section{References}
\label{sec:seq66_references}

   The \textsl{Sequencer66} reference list.

\begin{thebibliography}{99}

   \bibitem{alsa}
   ALSA team.
   \emph{Advanced Linux Sound Architecture (ALSA) project homepage.}
   \url{http://www.alsa-project.org/}.
   ALSA tools through version 1.0.29.
   2015.

   \bibitem{amsynth}
   amSynth team, Nick Dowell.
   \emph{amSynth and Demos with Calf Effects.}
   \url{http://amsynth.com/amsynth.html}.
   Includes links to demos and the source code.
   2015.

   \bibitem{bristol}
   Bristol team, Nick Copeland.
   \emph{Bristol: A Vintage Synthesizer Emulator.}
   \url{http://www.linuxsynths.com/BristolPatchesDemos/bristol.html}.
   2014.

   \bibitem{midimapper}
   Coolsoft.
   \emph{Coolsoft MIDIMapper (Windows).}
   \url{https://coolsoft.altervista.org/en/midimapper}.
   2018.

   \bibitem{midisynth}
   Coolsoft.
   \emph{Coolsoft VirtualMIDISynth (Windows).}
   \url{https://coolsoft.altervista.org/en/virtualmidisynth}.
   2018.

   \bibitem{fluidsynth}
   FluidSynth team.
   \emph{FluidSynth: A SoundFont Synthesizer.}
   \url{http://www.fluidsynth.org/}.
   2014.

   \bibitem{combine}
   Jay Capela Music.
   \emph{"Combine": A Seq24 Demonstration.}
   \url{https://www.youtube.com/watch?v=fUiXbVT0bJQ}.
   2010.

   \bibitem{jack}
   JACK team.
   \emph{JACK Audio Connection Kit.}
   \url{http://jackaudio.org/}.
   2015.

   \bibitem{kepler34}
   Oli Kester.
   \emph{Kepler34: Seq24 for the 2010s.}
   \url{https://github.com/oli-kester/kepler34}.
   2010-2016.

   \bibitem{layk}
   Lassi Ylikojola.
   \emph{Many demo videos of Sequencer66}
   \url{https://www.youtube.com/watch?v=YStYVjFv1TM},
   \url{https://www.youtube.com/watch?v=GBlEP8Ffqss},
   \url{https://www.youtube.com/watch?v=4gG8SvJxJkA&t=28s},
   \url{https://www.youtube.com/watch?v=n4Z4WPK6FpA}.
   2010-2017.

   \bibitem{linuxsynths}
   LinuxSynths team, briandc@linuxsynths.com.
   \emph{A Sonic Palette on the Linux Platform.}
   \url{http://www.linuxsynths.com/}.
   2015.

   \bibitem{ljtutorial}
   Dave Phillips.
   \emph{At the Sounding Edge: Introducing seq24.}
   \url{http://www.linuxjournal.com/article/8304}.
   Linux Journal, May 12, 2005.

   \bibitem{midicvt}
   Chris Ahlstrom.
   \emph{Extension of midicomp/midi2text to convert between MIDI and ASCII
      text format.}
   \url{https://github.com/ahlstromcj/midicvt}.
   2015-2016.

   \bibitem{midicontrol}
   linuxaudio.org.
   \emph{seq24: toggle sequences with a MIDI controller.}
   \url{http://wiki.linuxaudio.org/wiki/seq24togglemiditutorial}.
   2013.

   \bibitem{midicontroltable}
   midi.org.
   \emph{Summary of MIDI Messages.}
   \url{https://www.midi.org/specifications/item/table-1-summary-of-midi-message#2}.
   Year unknown.

   \bibitem{nanobasket}
   Roy Vegard.
   \emph{Configurator software for the Korg nanoSERIES of MIDI controllers.}
   \url{https://github.com/royvegard/Nano-Basket}.
   2015.

   \bibitem{portmidi}
   PortMedia team.
   \emph{Platform Independent Library for MIDI I/O.}
   \url{http://portmedia.sourceforge.net/portmidi/}.
   2010.

   \bibitem{rtmidi}
   Gary P. Scavone.
   \emph{The RtMIDI Tutorial.}
   \url{https://www.music.mcgill.ca/~gary/rtmidi/}.
   2016.

   \bibitem{seq24}
   Seq24 Team.
   \emph{The home site for the Sequencer66 looping sequencer.}
   \url{http://www.filter24.org/seq24/download.html}.
   2010.

   \bibitem{seq24demo1}
   pneumanlsd.
   \emph{Linux audio demo: Live sequencing with seq24.}
   \url{https://www.youtube.com/watch?v=f8zLV0vlSpY}.
   2010.

   \bibitem{seq24demo2}
   synthWF.
   \emph{Misty Corridor - Seq24 (with QSynth).}
   \url{https://www.youtube.com/watch?v=RH99zHvffGQ}.
   2012.

   \bibitem{seq24part1}
   pneumanlsd.
   \emph{Linux music tutorial: seq24, part 1.}
   \url{https://www.youtube.com/watch?v=J2WDHS1wYeM}.
   2010.
   
   \bibitem{seq24part2}
   pneumanlsd.
   \emph{Linux music tutorial: seq24, part 2.}
   \url{https://www.youtube.com/watch?v=i3Vpi3oxdqk}.
   2010.

   \bibitem{seq24launchpad}
   Seq24 Team.
   \emph{The home site for the Sequencer66 looping sequencer.}
   \url{https://launchpad.net/seq24}.
   2016.

   \bibitem{seq24launchpadmapper}
   Excds.
   \emph{A simple mapping for toggling the LEDs on the Novation launchpad
   together with seq24.}
   \url{https://github.com/Excds/seq24-launchpad-mapper}.
   2013.

   \bibitem{seq32}
   Stan Preston (stazed).
   \emph{The home site for the Seq32 looping sequencer.}
   \url{https://github.com/Stazed/seq32}.
   2016.

   \bibitem{seq42}
   sbrauer.
   \emph{The home site for the original Seq42 looping sequencer.}
   \url{https://github.com/sbrauer/seq42}.
   2016.

   \bibitem{seq42fork}
   Stan Preston (stazed).
   \emph{A fork of the Seq42 looping sequencer.}
   \url{https://github.com/Stazed/seq42}.
   2016.

   \bibitem{sequencer66}
   Chris Ahlstrom.
   \emph{A reboot of the Seq24 project as "Sequencer66".}
   \url{https://github.com/ahlstromcj/sequencer66/}.
   2015-2017.

   \bibitem{sequencer66doc}
   Chris Ahlstrom.
   \emph{The Sequencer66 User Manual.}
   \url{https://github.com/ahlstromcj/sequencer66-doc/}.
   2015-2017.

   \bibitem{sequencer66packages}
   Chris Ahlstrom.
   \emph{Sequencer66 Debian Packages.}
   \url{https://github.com/ahlstromcj/sequencer66-packages/}.
   2017.

   \bibitem{subatomicglue}
   Kevin at subatomicglue.com.
   \emph{Subatomic Mods for Seq24 Win32.}
   \url{http://www.subatomicglue.com/seq24/}.
   2010.

   \bibitem{timidity}
   Timidity++ Team.
   \emph{Download site for Timidity++ source code.}
   \url{http://sourceforge.net/projects/timidity/}.
   2015.

   \bibitem{vmpk}
   VMPK Team.
   \emph{Virtual MIDI Piano Keyboard.}
   \url{http://vmpk.sourceforge.net/}.
   2015.

   \bibitem{wootangent1}
   The Wootangent man.
   \emph{Sequencer66 Tutorial Video, Part 1.}
   \url{http://wootangent.net/2010/10/linux-music-tutorial-seq24-part-1/}.
   2010.

   \bibitem{wootangent2}
   The Wootangent man.
   \emph{Sequencer66 Tutorial Video, Part 2.}
   \url{http://wootangent.net/2010/10/linux-music-tutorial-seq24-part-2/}.
   2010.

   \bibitem{yoshimi}
   Yoshimi team \url{abrolag@users.sourceforge.net}.
   \emph{The download site for the Yoshimi software synthesizer.}
   \url{http://yoshimi.sourceforge.net/}.
   2015.

   \bibitem{yoshimi2}
   Yoshimi team.
   \emph{The alternate location for the Yoshimi source-code.}
   \url{https://github.com/abrolag/yoshimi/}.
   2015.

   \bibitem{yoshimidoc}
   Chris Ahlstrom.
   \emph{A Yoshimi User Manual.}
   \url{https://github.com/ahlstromcj/yoshimi-doc/}.
   2015-2017.

   \bibitem{yoshimicook}
   Chris Ahlstrom.
   \emph{A Yoshimi Cookbook.}
   \url{https://github.com/ahlstromcj/yoshimi-cookbook/}.
   2015.

   \bibitem{zynaddsubfx}
   Mark McCurry, Paul Nasca (ZynAddSubFX team).
   \emph{The download site for the ZynAddSubFX software synthesizer.}
   \url{http://zynaddsubfx.sourceforge.net/}.
   2015.

\end{thebibliography}

%-------------------------------------------------------------------------------
% vim: ts=3 sw=3 et ft=tex
%-------------------------------------------------------------------------------
